Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2020-03-03 xrb/submit/paper.tex
Migrate Don's changes over from parent dir
Sorry for the confusion

Creating the model involves specifying $T_\star$, $T_\mathrm{hi}$,
$T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
$\delta_\mathrm{atm}$, ${X_k}_\star$, ${X_k}_\mathrm{atm}$, and $g$.
We then integrate outwards from the base of the ramp region ($z =
H_\star$), enforcing the discrete form of hydrostatic equilibrium,
Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and $\rho_i$
using a Newton-Raphson solver together with the equation of state,
with either the temperature specified, $T_i = T(p_i, \rho_i,
\{X_k\}_i)$ (in the isothermal, ramp, and buffer layers) or constant
entropy, $s_i = s(p_i, \rho_i, \{X_k\}_i)$, in the fuel layer.  This
follows the procedures described in \citet{ppm-hse}.  We use a
constant temperature for all $z < H_\star + 3\delta_\mathrm{atm}$.
Above this, we switch to an isentropic atmosphere until the
temperature drops to \deleted{a floor value, }$T_\mathrm{lo}$, at
which point we again keep the temperature constant.  The integration
of the atmosphere continues until the density falls to a low density
cutoff, $\rho_\mathrm{cutoff}$.  The material above this height is
taken to have constant density and temperature.  The choice of factors
in front of $\delta_\mathrm{atm}$ were designed to make sure the peak
$T$ is attained at the desired density of the burning layer.  The
parameters we use for the model generation are listed in
Table~\ref{table:params} and the initial model profiles are showing in
Figure~\ref{fig:initial_models}.

@@

We also considered the effect of the network size.
Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
with the standard 13-isotope {\tt aprox13} network and the reduced
7-isotope {\tt iso7} network.  We see that the flame in the {\tt aprox13}
case is slightly more advanced and has a higher $\bar{A}$ than the {\tt iso7}
case.  We see in the next section that these two networks give largely
the same flame speed.  The reduced network size saves a lot of memory,
which will be useful when we transition to 3D simulations.

@@

All of the various approximations have the expected effects. The flame in the 5/5
boost run propagates at about half the speed of the 10/10 flame. This is consistent
with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
1000~Hz run goes \added{nearly} 2 times faster than the 10/10 2000~Hz run, as expected from the
inverse relation between rotation rate and the ratio of burning front area to scale height
\replaced{\citep{cavecchi:2013}}{(and as observed in \citealt{cavecchi:2013})}.
This confirms that the role of the Coriolis
force is to limit the rate of flame spreading by the anticipated geometrical/hydrodynamical
effect ($s_\mathrm{front} = s_L * L_R / H$). Reducing the resolution
had minimal impact on the flame speed -- the low resolution line is right on top of the fit for the
standard run in Figure \ref{fig:flame_speed}, with their slopes differing by only a few percent. Using a
smaller network also produces similar behavior to the standard run, although there is a small
reduction in speed owing to less energetic burning.

@@

In addition to extending our models to 3D and relaxing the
approximations used in this study, in the future we plan to perform a
number of additional further studies.  These include investigating the
effects of ignition latitude and different initial models. We also
plan to model mixed H/He bursts. This will require a different
reaction network, and perhaps different resolution
requirements. A long term goal is to include magnetohydrodynamics in
our models. Although the magnetic fields of neutron stars exhibiting
X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9$G
\citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
which have $B\lesssim 10^{15}$G), it has been found by
\citet{art-2016-cavecchi-etal} that even weak magnetic fields could
have a non-negligible effect on the flame propagation, reducing
confinement due to the Coriolis force and leading to increased flame
speeds. Ultimately, we wish to link our simulations to observed light
curves. To do this, we will need to explore radiation transport in
order to model how the burst energy propagates through the outer
layers of the neutron star atmosphere.

>>>>

Creating the model involves specifying $T_\star$, $T_\mathrm{hi}$,
$T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
$\delta_\mathrm{atm}$, ${X_k}_\star$, ${X_k}_\mathrm{atm}$, and $g$.
We then integrate outwards from the base of the ramp region ($z =
H_\star$), enforcing the discrete form of hydrostatic equilibrium,
Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and $\rho_i$
using a Newton-Raphson solver together with the equation of state,
with either the temperature specified, $T_i = T(p_i, \rho_i,
\{X_k\}_i)$ (in the neutron star, ramp, and isothermal buffer layers) or constant
entropy, $s_i = s(p_i, \rho_i, \{X_k\}_i)$, in the fuel layer.  This
follows the procedures described in \citet{ppm-hse}.  We use a
constant temperature for all $z < H_\star + 3\delta_\mathrm{atm}$.
Above this, we switch to an isentropic atmosphere until the
temperature drops to \deleted{a floor value, }$T_\mathrm{lo}$, at
which point we again keep the temperature constant.  The integration
of the atmosphere continues until the density falls to a low density
cutoff, $\rho_\mathrm{cutoff}$.  The material above this height is
taken to have constant density and temperature.  The choice of factors
in front of $\delta_\mathrm{atm}$ were designed to make sure the peak
$T$ is attained at the desired density of the burning layer.  The
parameters we use for the model generation are listed in
Table~\ref{table:params} and the initial model profiles are showing in
Figure~\ref{fig:initial_models}.

@@

We also considered the effect of the network size.
Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
with the standard 13-isotope {\tt aprox13} network and the reduced
7-isotope {\tt iso7} network.  We see that the flame in the {\tt aprox13}
case is slightly more advanced and has a higher $\bar{A}$ than the {\tt iso7}
case.  We see in the next section that these two networks give largely
the same flame speed.  We only ran the simulation with the {\tt iso7}
reaction network to 15~ms, and we observe this was sufficient to establish
a steady flame speed for comparison to our 10/10 simulation with the {\tt aprox13}
network. The reduced network size saves a lot of memory, which will be useful when we
transition to 3D simulations.

@@

All of the various approximations have the expected effects. The flame in the 5/5
boost run propagates at about half the speed of the 10/10 flame. This is consistent
with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
1000~Hz run goes nearly 2 times faster than the 10/10 2000~Hz run, as expected from the
inverse relation between rotation rate and the ratio of burning front area to scale height
\replaced{\citep{cavecchi:2013}}{(and as observed in \citealt{cavecchi:2013})}.
This confirms that the role of the Coriolis
force is to limit the rate of flame spreading by the anticipated geometrical/hydrodynamical
effect ($s_\mathrm{front} = s_L * L_R / H$). Reducing the resolution
had minimal impact on the flame speed -- the low resolution line is right on top of the fit for the
standard run in Figure \ref{fig:flame_speed}, with their slopes differing by only a few percent. Using a
smaller network also produces similar behavior to the standard run, although there is a small
reduction in speed owing to less energetic burning.

@@

In addition to extending our models to 3D and relaxing the
approximations used in this study, in the future we plan to perform a
number of additional further studies.  These include investigating the
effects of ignition latitude and different initial models. We also
plan to model mixed H/He bursts. This will require a different
reaction network, and perhaps different resolution
requirements. A long term goal is to include magnetohydrodynamics in
our models. Although the magnetic fields of neutron stars exhibiting
X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9~$G
\citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
which have $B\lesssim 10^{15}~$G), it has been found by
\citet{art-2016-cavecchi-etal} that even weak magnetic fields could
have a non-negligible effect on the flame propagation, reducing
confinement due to the Coriolis force and leading to increased flame
speeds. Ultimately, we wish to link our simulations to observed light
curves. To do this, we will need to explore radiation transport in
order to model how the burst energy propagates through the outer
layers of the neutron star atmosphere.
